Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-05 paper/paper.tex
pass on 4.3.2 and 4.3.3

\subsubsection{Popularity of options.}
One heuristic function that \tname{} uses to measure similarity
between a synthesized rule and a human-created rule is the popularity
of the options used in those rules. According to this heuristic, rules
containing popular options score higher compared to those containing
unpopular rules. Figure~\ref{fig:distribution-options} shows the
histogram of rule options for the ruleset we
used~\cite{emerging-threats-open}. The x-axis shows the options used
in rules (the 10 most common) and the y-axis shows the number of rules
using the corresponding option. Note that the options
\CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce} are very
popular. \tname\ uses the following formula to compute the
contribution of option popularity to the score of a rule $r$:

\vspace{-2ex}
\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]
\vspace{-2ex}

@@

\Mar{------------------------------------------ estou aqui}
\subsubsection{Number of content options.} We observed
that \percContentOptions\ of the rules from the ruleset we are
analyzing contain content options, but the distribution of those
contents per rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
corresponding distribution. Given the importance of this kind of
option to the definition of rules, we defined a heuristic function
focused on the typical number of content options in a rule. We
calculate the contribution of this function as described on previous
Section~\ref{sec:heuristic-size-of-rule}.

%% to the score of a rule
%% with the following formula:

%% \vspace{-2ex}
%% \[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
%% \vspace{-2ex}

@@

\subsubsection{Popularity of content strings.} 
The heuristic described in this section leverages content rarity to
discriminate rules---it weighs higher (respectively, lower) rules that
use popular (respectively, unpopular) strings in content options. The
score of a rule $r$ on this function is given by the arithmetic mean
of the popularity scores of each string used in contents in $r$. More
precisely, the contribution of content rarity is given by the
following formula:

\vspace{-2.5ex}
\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
\vspace{-2.5ex}

\noindent
, where $\mathit{content_i}$ denotes the $i$-th content in a given
rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our ruleset. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2). The relatively high weight reflects the fact that
``GET'' and ``ID'' are popular contents.

>>>>

\subsubsection{\label{sec:popularity-options}Popularity of options.}
One heuristic function that \tname{} uses to measure similarity
between a synthesized rule and a human-created rule is the popularity
of the options used in those rules. According to this heuristic, rules
containing popular options score higher compared to those containing
unpopular rules. Figure~\ref{fig:distribution-options} shows the
histogram of rule options for the ruleset we
used~\cite{emerging-threats-open}. The x-axis shows the options used
in rules (the 10 most common) and the y-axis shows the number of rules
using the corresponding option. Note that the options
\CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce} are very
popular. \tname\ uses the following formula to compute the
contribution of option popularity to the score of a rule $r$:

\vspace{-2ex}
\[\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N\]
\vspace{-2ex}

@@

\subsubsection{Number of content options.} We observed
that \percContentOptions\ of the rules from the ruleset we analyzed
contain content options, but the distribution of those contents per
rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
distribution of number of contents in a rule. Results indicate nearly
73\% of the rules from our ruleset contain one to three content
options. Hence, we found that that information could help
\tname\ discriminate useful rules. \tname{} computes the contribution
of the number of content options to the score of a rule $r$ with the
fraction
$\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}$,
where $\mathit{len(r)}$ denotes the number of content options in $r$
and $\mathit{rules\_per\_number\_of\_contents}$ denotes the
distribution of number of contents (encoded with a dictionary as in
the function described on
Section~\ref{sec:popularity-options}). Consequently the term
$\mathit{rules\_per\_number\_of\_contents[len(r)]}$ corresponds to the
number of rules in the ruleset with the same number of contents as
$r$. The denominator of the fraction normalizes the contribution in
the 0-1 range. To illustrate, the contribution of this heuristic
function to the score a rule would be $.64$ $(=5841/9164)$ to a rule
with 3 content options and $.18$ $(=1622/9164)$ to a rule with 5
content options.

@@

\subsubsection{Popularity of content strings.} 
This heuristic leverages content rarity to discriminate rules---it
weighs higher (respectively, lower) rules that use popular
(respectively, unpopular) strings in content options. The score of a
rule $r$ on this function is given by the arithmetic mean of the
popularity scores of each string used in contents in $r$. More
precisely, the contribution of content rarity is given by the
following formula:

\vspace{-1ex}
\[\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}/N\]
\vspace{-1ex}

\noindent
, where $\mathit{content_i}$ denotes the $i$th content in a given
rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our ruleset. To illustrate, let us
consider a rule with whose content options use the strings ``GET'' and
``ID'', respectively. The contribution of that rule for the
computation of similarity score would be 0.601 (=(1+.203)/2). The
relatively high weight reflects the fact that ``GET'' and ``ID'' are
popular content strings.

\Mar{------------------------------------------------ estou aqui}
